Robust receding horizon control for linear systems with state dependent disturbance

### 2015-04-21 example.tex
Preliminary synch.

\section{Example}\label{sec:example}
This section we present a numerical simulation in which the state-dependent uncertainty enters through an upper bound on the linearisation
error for a nonlinear system. We consider an inverted pendulum described by $ml^2\ddot\varphi = mlg\sin\varphi+M$, where
$m=0.1$\,kg is the mass, $l=0.3$\,m is the length of the pendulum, $g=9.81$\,m\,s$\mbox{}^{-2}$ is the gravitational
constant and $M$ is the direct torque input. Discretisation using Euler backward finite differences 
at $f = 20$\,Hz yields
\begin{equation}
  \begin{split}
    x_1[k+1] &= x_1[k] + \frac{1}{f} x_2[k]\\
    x_2[k+1] &= \frac{g}{fl} \sin(x_1[k]) + x_2[k] + \frac{1}{fml^2}u[k]
  \end{split}
\end{equation}
where $x_1[k] = \varphi(t_0+k/f),x_2[k] = \dot\varphi(t_0+k/f)$ and $u[k] = M(t_0+k/f)$.
The nonlinearity is given by the sine function, and we can explicitly state the 
linearisation error as $e(x_1)=\frac{g}{fl}(\sin x_1 - x_1)$. Since the linearisation
only approximates the nonlinear dynamics around the equilibrium we can choose a closed interval 
$0\in D\subset\mathbb R$ and maximise the upper bound 
\begin{equation}
  \Delta e = \max_{x_1\in D}\abs{\frac{de}{dx_1}} = \frac{g}{fl}\max_{x_1\in D}\abs{\cos x_1-1}.
\end{equation}
This results in the linear system
\begin{equation}
  x^+ = \underbrace{\left(\begin{array}{cc}
  1 & f^{-1}\\ g(fl)^{-1} & 1
  \end{array}\right)}_A x + \underbrace{\left(\begin{array}{c} 0 \\(ml^2f)^{-1} \end{array}\right)}_B u
  +\left(\begin{array}{c}w_1\\ w_2\end{array}\right)
\end{equation}
with $\abs{w_1}\leq w_{1,\max}$ and $\abs{w_2}\leq\max\{w_{2,\max},\Delta e \abs{x_1}\}$.
We also introduce the input constraint $\abs{u}\leq3$\,N\,m and use the interval $D = [-\frac{3\pi}{4},
\frac{3\pi}{4}]$ and a horizon of $N=7$ to obtain the sequence of state constraints $\mathcal X_m$
depicted in Figure~\ref{fig:numerical:example}. The line search starting at the origin and
exploring towards $\mathpzc x_e=(-6,50)$ produces the optimal trajectories at the active-set switches
which are illustrated in blue. In pink a warm-started line search terminates on the boundary of the
feasible set before reaching the (infeasible) state $\mathpzc x_e=(0,30)$.

>>>>

%!TEX root = draft_v2.tex
\section{Example}\label{sec:example}
In this section we present a numerical simulation for a simplified levitating ball model, as derived in 
by~\cite{Schaich:2015}. The dynamics of the nonlinear system are given $m\ddot y = m g - c\frac{i^2}{y^2}$,
where $m$ denotes the mass of the ball, $g$ denotes the gravitation constant, $c$ denotes a constant depending
on the coil, $y$ the vertical distance to the coil, the simplification is due to neglecting inductive dynamics 
and assuming we that can influence the current~$i$ directly. Figure~\ref{fig:levitating:ball} illustrates the setup.
Choosing the state $x=(y,\dot y)$ and the input $u=i$ we obtain a state space representation which depends nonlinearly
on both the state and the input. An equilibrium is given for all $u=\sqrt{\frac{gm}{c}}x_1$.

Linearising the nonlinear differential 
equation $\dot x = f(x,u)$ around an equilibrium point $(\hat x, \hat
u)$ gives the approximate linear model 
%
\[
	 \Delta\dot{x} = \underbrace{\left(\begin{array}{cc}
	0 & 1 \\ \frac{2c\hat u^2}{m\hat x_1^3} & 0
	\end{array}\right)}_{\frac{\partial f}{\partial x}(\hat x,\hat
      u)}\Delta x 
+ \underbrace{\left(\begin{array}{c}
	0 \\ - \frac{2c\hat u}{m\hat x_1^2}
	\end{array}\right)}_{\frac{\partial f}{\partial u}(\hat x,\hat
      u)}\Delta u
\]
%
where $\Delta u = u -\hat{u}$ and $\Delta x \approx x-\hat{x}$.
We derive the discrete time dynamics with sampling rate $T_s$ using
the Euler formula $x^+=x+T_s f(x,u) =:\tilde f(x,u)$ giving
\[
\Delta x^+ = A \Delta x + B \Delta u , \quad
\Biggl\{\begin{aligned} A &= I+T_s\frac{\partial f}{\partial  x}(\hat
  x,\hat u) \\
B &= T_s \frac{\partial f}{\partial u}(\hat x,\hat u)
\end{aligned}
\]
%
A representation of the form~\eqref{eq:the:problem}

\begin{figure}
\centering
% \begin{overpic}[scale=0.75]{levitatingBall}
% \put(30,5){$m g$}
% \put(30,41){$c\frac{i^2}{x^2}$}
% \put(78,28){$y$}
% \put(90,89){$i$}
% \end{overpic}
\begin{lpic}{levitatingBall(.75,)}
\lbl[tr]{25,3; $m g$}
\lbl[br]{25,25; $c\frac{i^2}{y^2}$}
\lbl[bl]{49,17; $y$}
\lbl[bl]{56,55; $i$}
\end{lpic}
\vspace{-2mm}
\caption{Levitating ball system.}
\label{fig:levitating:ball}
\end{figure}

% This section we present a numerical simulation in which the state-dependent uncertainty enters through an upper bound on the linearisation
% error for a nonlinear system. We consider an inverted pendulum described by $ml^2\ddot\varphi = mlg\sin\varphi+M$, where
% $m=0.1$\,kg is the mass, $l=0.3$\,m is the length of the pendulum, $g=9.81$\,m\,s$\mbox{}^{-2}$ is the gravitational
% constant and $M$ is the direct torque input. Discretisation using Euler backward finite differences 
% at $f = 20$\,Hz yields
% \begin{equation}
%   \begin{split}
%     x_1[k+1] &= x_1[k] + \frac{1}{f} x_2[k]\\
%     x_2[k+1] &= \frac{g}{fl} \sin(x_1[k]) + x_2[k] + \frac{1}{fml^2}u[k]
%   \end{split}
% \end{equation}
% where $x_1[k] = \varphi(t_0+k/f),x_2[k] = \dot\varphi(t_0+k/f)$ and $u[k] = M(t_0+k/f)$.
% The nonlinearity is given by the sine function, and we can explicitly state the 
% linearisation error as $e(x_1)=\frac{g}{fl}(\sin x_1 - x_1)$. Since the linearisation
% only approximates the nonlinear dynamics around the equilibrium we can choose a closed interval 
% $0\in D\subset\mathbb R$ and maximise the upper bound 
% \begin{equation}
%   \Delta e = \max_{x_1\in D}\abs{\frac{de}{dx_1}} = \frac{g}{fl}\max_{x_1\in D}\abs{\cos x_1-1}.
% \end{equation}
% This results in the linear system
% \begin{equation}
%   x^+ = \underbrace{\left(\begin{array}{cc}
%   1 & f^{-1}\\ g(fl)^{-1} & 1
%   \end{array}\right)}_A x + \underbrace{\left(\begin{array}{c} 0 \\(ml^2f)^{-1} \end{array}\right)}_B u
%   +\left(\begin{array}{c}w_1\\ w_2\end{array}\right)
% \end{equation}
% with $\abs{w_1}\leq w_{1,\max}$ and $\abs{w_2}\leq\max\{w_{2,\max},\Delta e \abs{x_1}\}$.
% We also introduce the input constraint $\abs{u}\leq3$\,N\,m and use the interval $D = [-\frac{3\pi}{4},
% \frac{3\pi}{4}]$ and a horizon of $N=7$ to obtain the sequence of state constraints $\mathcal X_m$
% depicted in Figure~\ref{fig:numerical:example}. The line search starting at the origin and
% exploring towards $\mathpzc x_e=(-6,50)$ produces the optimal trajectories at the active-set switches
% which are illustrated in blue. In pink a warm-started line search terminates on the boundary of the
% feasible set before reaching the (infeasible) state $\mathpzc x_e=(0,30)$.
